Refinement of Thermostated Molecular Dynamics Using Backward Error Analysis

### 2018-11-11 rigid_bodies.tex
Further changes in discretization effects on kinetic energies

We end this section with a discussion about discretization effects on the kinetic energy partition amongst translational and rotational degrees of freedom.
Davidchack \cite{Davidchack_2010} has already performed a thorough analysis of such effects in thermostated rigid-body dynamics.
Here we complement his study by including the KLN method \cite{Kamberaj_2005} and by showing that the refinement procedure is able to mitigate the observed artifacts.
In Fig.~\ref{fig:energy_partition}, we present computed averages of the translational ($K_t$) and rotational ($K_r$) kinetic energies as functions of the time-step size, and compare them to the values that would be expected by assuming equipartition.
In Figs.~\ref{fig:energy_partition}(a) to \ref{fig:energy_partition}(c), these were calculated from the average kinetic energy $\langle K \rangle$ by means of $\langle K_t \rangle_\text{eq}$ = $\frac{3N-3}{6N-3} \langle K \rangle$ and $\langle K_r \rangle_\text{eq}$ = $\frac{3N}{6N-3} \langle K \rangle$.
For Fig.~\ref{fig:energy_partition}(d), $K$ is simply replaced by $\refined{K}$ in these equations.
In a previous work \cite{Silveira_2017}, we have shown that the rotational energy is more strongly influenced by the time-step size in the NVE case than the translational energy.
This occurs because the rotational degrees of freedom display the fastest motion in a rigid water molecule \cite{Silveira_2017}.
In Fig.~\ref{fig:energy_partition}(a), one can see that the KLN method \cite{Kamberaj_2005} tends, once again, to reproduce the behavior of the NVE integration.
The results of Fig.~\ref{fig:energy_partition}(b) were obtained by using the NHC scheme of Martyna \textit{et al}. \cite{Martyna_1996} applied to rigid bodies.
These results are similar to those shown in Fig.~\ref{fig:energy_partition}(a), in the sense that the simulated translational and rotational components deviate upwards and downwards, respectively, from their equipartioned values.
This observation is in consonance with the analysis of Ref.~\citenum{Davidchack_2010}.
To some extent, our refinement/reweighting procedure is able to correct these deviations, as one can see in Fig.~\ref{fig:energy_partition}(c).
Finally, by observing Fig.~\ref{fig:energy_partition}(d), one can note that the refined translational and rotational kinetic energies, being estimates of their ``shadow'' counterparts, comply with Tolman's equipartition theorem \cite{Tolman_1918} for time-step sizes up to $5~\text{fs}$.

>>>>

We end this section with a discussion about discretization effects on the kinetic energy partition amongst translational and rotational degrees of freedom.
Davidchack \cite{Davidchack_2010} has already performed a thorough analysis of such effects in thermostated rigid-body dynamics.
Here we complement his study by including the KLN method \cite{Kamberaj_2005} and by showing that the refinement procedure is able to mitigate the observed artifacts.
In Fig.~\ref{fig:energy_partition}, we present computed averages of the translational ($K_t$) and rotational ($K_r$) kinetic energies as functions of the time-step size, and compare them to the values that would be expected by assuming equipartition.
In Figs.~\ref{fig:energy_partition}(a) to \ref{fig:energy_partition}(c), these were calculated from the average kinetic energy $\langle K \rangle$ by means of $\langle K_t \rangle_\text{eq}$ = $\frac{3N-3}{6N-3} \langle K \rangle$ and $\langle K_r \rangle_\text{eq}$ = $\frac{3N}{6N-3} \langle K \rangle$.
For Fig.~\ref{fig:energy_partition}(d), $K$ is simply replaced by $\refined{K}$ in these equations.
In a previous work \cite{Silveira_2017}, we have shown that the rotational energy is more strongly influenced by the time-step size in the NVE case than the translational energy.
This occurs because the rotational degrees of freedom display the fastest motion in a rigid water molecule \cite{Silveira_2017}.
In Fig.~\ref{fig:energy_partition}(a), one can see that the KLN method \cite{Kamberaj_2005} tends, once again, to reproduce the behavior of the NVE integration.
The results of Fig.~\ref{fig:energy_partition}(b) were obtained by using the NHC scheme of Martyna \textit{et al}. \cite{Martyna_1996} applied to rigid bodies.
These results are similar to those shown in Fig.~\ref{fig:energy_partition}(a), in the sense that the simulated translational and rotational components deviate upwards and downwards, respectively, from their equipartioned values.
This observation is in consonance with the analysis of Ref.~\citenum{Davidchack_2010}.
To some extent, our refinement/reweighting procedure is able to correct these deviations, as one can see in Fig.~\ref{fig:energy_partition}(c).
Finally, by observing Fig.~\ref{fig:energy_partition}(d), one can note that the refined translational and rotational kinetic energies, 
$\refined K_t = \frac{1}{2} \tr{\vt p} \dot{\vt r}^{[3]}$ and $\refined K_r = \frac{1}{2} \tr{\vt \pi} \dot{\vt q}^{[3]}$, being estimates of their ``shadow'' counterparts, comply with Tolman's equipartition theorem \cite{Tolman_1918} for time-step sizes up to $5~\text{fs}$.
